Reformulating phylogenetic mixed models to improve flexibility and speed

### 2020-02-11 phyloglmm_ms.tex
cutted down to 3300 words

Ever-increasing data collection capabilities, in combination with large-scale synthetic databases of species occurrence and phenotypic traits, are making larger volumes of biological data available over an ever-wider taxonomic range.
Researchers use these data to fit complex models describing species occurrence and traits \citep{garland1992procedures, freckleton2002phylogenetic, ord2010adaptation, davies2013phylogenetic, winter2013phylogenetic, santamaria2012evolution, lankau2011incorporating, lavergne2010biodiversity, mace2008evolutionary}.

@@

Classic phylogenetic regressions usually allow the uncertainty in the response (trait or distribution) to evolve according to the phylogenetic relationship across species, but the effects of the predictor variables may evolve according to the phylogenetic relationship across species as well. 
For example, suppose we have examined a collection of species that came from two groups, and wish to know whether their brain size (Y) is proportional to their body size (X) (Felsenstein's (1985) example using a mixed-effect model. 
The standard phylogenetic mixed-effect model allows for phylogenetic correlations in the intercept of the relationship between body and brain size. 
However, species within taxonomic groups with similar body sizes may vary in brain size (random intercepts), or taxonomic groups may vary in the relationship between brain and body size (random slopes). 
Analogous to phylogenetically correlated variation in intercepts, phylogenetic random slopes allow the relationship between predictors and responses to vary between species in a phylogenetically correlated way.
Several recent studies have incorporated some form of phylogenetic variation using random effects in addition to the standard phylogenetic correlations in the intercepts and looked at species response to phylogenetic variation with changes in environmental factors \cite{nowakowski2018phylogenetic, li2017canfun}.
The tools available for extending phylogenetic relationships to predictor variables (or predictor level variation) in a standard frequentist framework are relatively inflexible; thus, biologists needing to fit random-slopes model typically turn to more flexible Bayesian approaches, despite their additional computational burden \citep{hadfield2010mcmc, burkner2018brms}.
Table \ref{table:model} summarizes currently available methods for phylogenetic comparative analysis.

@@

We used our method to reproduce the examples in chapter 11 of \cite{garamszegi2014modern} using phylogenetic GLMMs based on \pkg{lme4} and \pkg{glmmTMB} (for more details see example in supplement).

>>>>

@@

Classic phylogenetic regressions usually allow the uncertainty in the response (trait or distribution) to evolve according to the phylogenetic relationship across species, but the effects of the predictor variables may evolve according to the phylogenetic relationship across species as well. 
For example, suppose we have examined a collection of species that came from two groups, and wish to know whether their brain size (Y) is proportional to their body size (X) (Felsenstein's (1985) example using a mixed-effect model. 
The standard phylogenetic mixed-effect model allows for phylogenetic correlations in the intercept of the relationship between body and brain size. 
However, species within taxonomic groups with similar body sizes may vary in brain size (random intercepts), or taxonomic groups may vary in the relationship between brain and body size (random slopes). 
Analogous to phylogenetically correlated variation in intercepts, phylogenetic random slopes allow the relationship between predictors and responses to vary between species in a phylogenetically correlated way.
Several recent studies have incorporated some form of phylogenetic variation using random effects in addition to the standard phylogenetic correlations in the intercepts and looked at species response to phylogenetic variation with changes in environmental factors \citep{nowakowski2018phylogenetic, li2017canfun}.
The tools available for extending phylogenetic relationships to predictor variables (or predictor level variation) in a standard frequentist framework are relatively inflexible; thus, biologists needing to fit random-slopes model typically turn to more flexible Bayesian approaches, despite their additional computational burden \citep{hadfield2010mcmc, burkner2018brms}.
Table \ref{table:model} summarizes currently available methods for phylogenetic comparative analysis.

@@

We used our method to reproduce the examples in chapter 11 of \cite{garamszegi2014modern} using phylogenetic GLMMs based on \pkg{lme4} and \pkg{glmmTMB} (for more details see Data Availability section).
